---
title: "STATS 700-002 Class 2.\\newline The coalescent"
author:
  - "Edward Ionides and Aaron King"
date: September 4, 2025
date-format: long
format:
  beamer:
    cite-method: natbib
    biblio-style: apalike
    incremental: true
    slide-level: 2
    classoption: t
    nocite: |
      @Kingman1982a, @Kingman1982b, @Kingman1982c, @Wakeley2009
    header-includes: |
      \setbeamertemplate{footline}[page number]

bibliography: ../phylodynamics.bib
---

## The coalescent

Kingman, J. F. C. (1982). The coalescent. _Stochastic Processes and their Applications_, 13(3), 235-248. [https://doi.org/10.1016/0304-4149(82)90011-4](https://doi.org/10.1016/0304-4149(82)90011-4).


## Section 1

- How does the construction of section 1 relate to a genealogical tree?

- What does the assumption (1.3) mean?

- Is the relation $\prec$ defined in (1.4) an order relation?

- How can we see that $q_\xi=\binom{|\xi|}{2}$, as claimed in (1.6)?

- In terms of genealogical trees, what is the interpretation of the $D_t$ process introduced in (1.8)?

- In terms of genealogical trees, what does the *transit time* $T$ introduced in (1.11) correspond to?

## Coalescent trees

```{r plots,echo=FALSE}
library(tidyverse)
library(phylopomp)
set.seed(5996843)
```

```{r,echo=FALSE}
simulate("Moran",time=30,n=20,psi=0) |>
  plot(prune=FALSE)+
  labs(x="time")
```

-----------------

```{r,echo=FALSE}
simulate("Moran",time=10,n=20,psi=0) |>
  plot(prune=FALSE)+
  labs(x="time")
```

## Theorem 1

- What do we have to know about continuous time Markov chains to follow this proof?

- Why does equation (2.1) hold?

- How can we see why equation (2.2) holds?


## Theorems 3 and 4

- What does Theorem 3 accomplish?

- What is accomplished by Theorem 4?

## Relevance to phylodynamics

- What is the relevance of Kingman's coalescent to practical phylodynamic inference?

- Given what we know at this point, how might Kingman's coalescent be used?

- What extensions can you envision?

# Bibliography
